Methods for Large Scale Learning to Rank

### 2014-04-28 report/Chapters/LTRMethods.tex
nesterov's accelerated gradient descent uitgelegd

\LinesNumbered
\begin{algorithm}[H]
 \KwData{pairwise training data matrix $K$}
 \KwIn{desired accuracy $\epsilon$, maximum number of iterations $T$ and the radius $r$ of the $\ell1$ ball.}
 Initialize: $w_1$ = $0_m$\\
 \For{$t\leftarrow 1$ \KwTo $T$}{
 	//check if the early stopping criterion is satisfied\\
 	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
 	//here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$, \\
 	// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and\\
 	// $g_{t} = d_{t}^{T}K$\\
 	return $w_{t}$ as ranking predictor $w$
 	}
 //greedily choose a feature to update\\
 Choose $j_{t} = \argmax_{j}|(g_t)_j)|$\\
 //compute an appropriate step size\\
 Let $\mu_t = \argmax_{0 \le \mu_{t} \le 1} D((1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})_{j_{t}})e^{j_{t}})$\\
 //update the model with the chosen feature and step size\\
 Update $w_{t+1}=(1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})j_{t})e^{j_{t}}$
 }
 return $w_{T}$ as ranking predictor for $w$
 \caption{Learning algorithm of FenchelRank, obtained from \cite{Lai2013}}
 \label{alg:fenchelrank}
\end{algorithm}
\section{FSMRank}
Lai et al. \cite{Lai2013c} observed that existing feature selection methods in Learning-to-Rank all follow a two-stage paradigm: 1) select subset of features from the original features, and 2) learn a ranking model based on the selected features. It is stated as a limitation of this paradigm that the selected features in the first step are not necessarily the optimal features for the second stage where the ranking model is build. FSMRank \cite{Lai2013c} addresses this limitation by using a joint convex optimisation problem that minimises ranking errors while simultaneously selecting a subset of features.
\LinesNumbered
\begin{algorithm}[H]
 \KwData{training set $S = {(q_k,\hat{X}^{(q_k)},Y^(q_k)}_{k=1}^{n}$}
 \KwIn{$\lambda_1, \lambda_2, T \text{ and } L_0$}
 Initialize: $w_0=z_1=0, \alpha_1=1, \gamma=2, L=L_0/\gamma^{10}$\\
 \For{$t\leftarrow 1$ \KwTo $T$}{
 	Let $g = l^{'}(z_t)$\\
 	\While{true}{
 		//projection step\\
 		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$ s\\
 		\If{$l(w_t) \le l(z_t) + \langle l^{'}(z_t),w_t-z_t \rangle +\frac{L}{2}||w_t-z_t||^2$}{
 			break\\
 		}
 		$L = \lambda L$ 		
 	}
 	\If{$\frac{|F(w_t)-F(w_{t-1})|}{|F(w_{t-1})|} \le \epsilon_s$}{
 		//early stopping criterion\\
 		break
 	}
 	$\alpha_{t+1}=\frac{1+\sqrt{1+4\alpha_t^2}}{2}$\\
 	$z_{t+1} = w_t + \frac{\alpha_t-1}{\alpha_t+1}(w_t-w_{t-1})$ 
 }

>>>>

\LinesNumbered
\begin{algorithm}[H]
 \KwData{pairwise training data matrix $K$}
 \KwIn{desired accuracy $\epsilon$, maximum number of iterations $T$ and the radius $r$ of the $\ell1$ ball.}
 Initialize: $w_1$ = $0_m$\\
 \For{$t\leftarrow 1$ \KwTo $T$}{
 	//check if the early stopping criterion is satisfied\\
 	\If{$||g_{t}||_{\infty} + \langle d_{t}, -Kw_{t} \rangle \le \epsilon$}{
 	//here $d_{t} = \nabla f^{*}(-Kw_{t}) = \frac{\partial f^{*}(-Kw)}{\partial(Kw)}|w=w_{t}$, \\
 	// $\langle x,y \rangle$ represents the inner products of vectors $x$ and $y$, and\\
 	// $g_{t} = d_{t}^{T}K$\\
 	return $w_{t}$ as ranking predictor $w$
 	}
 //greedily choose a feature to update\\
 Choose $j_{t} = \argmax_{j}|(g_t)_j)|$\\
 //compute an appropriate step size\\
 Let $\mu_t = \argmax_{0 \le \mu_{t} \le 1} D((1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})_{j_{t}})e^{j_{t}})$\\
 //update the model with the chosen feature and step size\\
 Update $w_{t+1}=(1-\mu_{t})w_{t} + \mu_{t}\sign((g_{t})j_{t})e^{j_{t}}$
 }
 return $w_{T}$ as ranking predictor for $w$
 \caption{Learning algorithm of FenchelRank, obtained from \cite{Lai2013}}
 \label{alg:fenchelrank}
\end{algorithm}
\section{FSMRank}
Lai et al. \cite{Lai2013c} observed that existing feature selection methods in Learning-to-Rank all follow a two-stage paradigm: 1) select subset of features from the original features, and 2) learn a ranking model based on the selected features. It is stated as a limitation of this paradigm that the selected features in the first step are not necessarily the optimal features for the second stage where the ranking model is build. FSMRank \cite{Lai2013c} addresses this limitation by using a joint convex optimisation problem that minimises ranking errors while simultaneously selecting a subset of features.\\

FSMRank uses an extended version of gradient descent optimisation, proposed by Yuri Nesterov, that enables faster convergence for convex problems \cite{Nesterov2004}. Nesterov's accelerated gradient descent can guarantee an $\epsilon$-accurate solution in at most $T$ iterations where $\epsilon=\mathcal{O}(1/T^2)$.

\LinesNumbered
\begin{algorithm}[H]
 \KwData{training set $S = {(q_k,\hat{X}^{(q_k)},Y^(q_k)}_{k=1}^{n}$}
 \KwIn{$\lambda_1, \lambda_2, T \text{ and } L_0$}
 Initialize: $w_0=z_1=0, \alpha_1=1, \gamma=2, L=L_0/\gamma^{10}$\\
 \For{$t\leftarrow 1$ \KwTo $T$}{
 	Let $g = l^{'}(z_t)$\\
 	\While{true}{
 		//projection step\\
 		$w_t= \argmin_{w_t:w_t\in \mathbb{R}_{+}^{2d}} Q(w_t,z_t)$ s\\
 		\If{$l(w_t) \le l(z_t) + \langle l^{'}(z_t),w_t-z_t \rangle +\frac{L}{2}||w_t-z_t||^2$}{
 			break\\
 		}
 		$L = \lambda L$ 		
 	}
 	\If{$\frac{|F(w_t)-F(w_{t-1})|}{|F(w_{t-1})|} \le \epsilon_s$}{
 		//early stopping criterion\\
 		break
 	}
 	$\alpha_{t+1}=\frac{1+\sqrt{1+4\alpha_t^2}}{2}$\\
 	$z_{t+1} = w_t + \frac{\alpha_t-1}{\alpha_t+1}(w_t-w_{t-1})$ 
 }
